---
title: "Exploring potential explanations for eligibility drift"
author: "Xing Song"
format: html
editor: visual
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
#| label: load-pkg

rm(list=ls())
pacman::p_load(
  tidyverse,
  magrittr,
  stringr,
  broom,
  cowplot,
  ggrepel,
  kableExtra,
  gtsummary,
  devtools,
  tableone,
  pROC
)
source_url("https://raw.githubusercontent.com/sxinger/utils/master/analysis_util.R")
source_url("https://raw.githubusercontent.com/sxinger/utils/master/plot_util.R")

path_to_data<-"C:/repos/r61-r33_vccc_kumc/data"
path_to_res<-"C:/repos/r61-r33_vccc_kumc/res"
```

## Clinic BP vs. Home BP at Baseline

There are 1000 in the baseline BP file, however, after matching with CDM tables, we missed 102 participants due to mapping issue in PCORnet_TRIAL table (missing a link between STUDY_ID and PATID in the PCORnet CDM data), and 10 participants due to missing DOB in the PCORnet CDM table. We ended up with 888 participants with eligible data.

Run-in period is defined as the time gap between eligibility clinic visit and baseline visit. We define the following stratification groups:

1.  based on eligibility sbp:

-   esbp1: 140 - 150
-   esbp2: 150 - 160
-   esbp3: \> 160

2.  based on baseline sbp:

-   bsbp1: 140 - 150
-   bsbp2: 150 - 160
-   bsbp3: \> 160

3.  based on change from eligibility to baseline sbp:

-   deltasbp1: sbp decreased to below 130
-   deltasbp2: sbp decreased to between 130 and 140
-   deltasbp3: sbp decreased but remain above 140 or not decreased

```{r}
#| label: bp-baseline-plot

base_bp<-readRDS(file.path(path_to_data,'base_bp.rda')) %>%
  semi_join(readRDS(file.path(path_to_data,'baseline_aset.rda')),by="STUDY_ID")

N<-length(unique(base_bp$STUDY_ID))
ggplot(
  base_bp %>%
    filter(ELIG_SINCE_INDEX>=-100&ELIG_SINCE_INDEX<=0) %>%
    select(STUDY_ID,ELIG_SINCE_INDEX) %>%
    unique,
  aes(x=ELIG_SINCE_INDEX)
)+
  geom_histogram(
    aes(y=..density..),
    binwidth = 2,
    fill="grey",
    color="blue")+
  geom_density(
    aes(y = ..density..),
    color = "red",
    stat = 'density',
    bw = 2,
    size = 1.5) +
  stat_bin(
    aes(y = ..density..,
        label=scales::percent(round(..density..,3))),
    geom='text',
    binwidth=2,
    vjust = 1.5,
    fontface = "bold"
  )+
  scale_x_continuous(
    name = "run-in period duration"
  )+
  scale_y_continuous(
    name = "Percentage (%)",
    breaks = seq(0,0.2,by=0.02),
    labels = seq(0,0.2,by=0.02)*100,
    sec.axis = sec_axis(
      trans = ~ . * N,
      name = "Count",
      breaks = seq(0,40000,by=2000),
      labels = seq(0,40000,by=2000))
  )+
  theme(
    panel.background = element_rect(fill = "white", colour = "grey50"),
    panel.grid.major.y = element_line(colour = "grey"),
    text = element_text(face="bold",size=14)
  )

ggplot(
  base_bp %>%
    select(STUDY_ID,ELIG_SINCE_INDEX,DELTA_SBP,DELTA_SBP_GROUP,SITE) %>%
    unique,
  aes(
    x = ELIG_SINCE_INDEX,
    y = DELTA_SBP,
    color = DELTA_SBP_GROUP
    )
)+
  geom_point()+
  geom_hline(yintercept=0,linetype=2)+
  geom_vline(xintercept=0,linetype=2)+
  geom_vline(xintercept=-30,linetype=2)+
  geom_vline(xintercept=-60,linetype=2)+
  geom_vline(xintercept=-90,linetype=2)+
  labs(
    x = 'run-in period duration', 
    y = 'SBP difference', 
    title = "eligibility and baseline SBP"
  )+
  facet_wrap(~ SITE)

elig_drift<-base_bp %>%
  group_by(SITE) %>%
  mutate(n_site = length(unique(STUDY_ID))) %>%
  ungroup %>%
  group_by(SITE,DELTA_SBP_GROUP,n_site) %>%
  summarise(cnt = length(unique(STUDY_ID)),.groups='drop') %>%
  mutate(
    N = n_site,
    prop = round(cnt/N,3)
  ) %>%
  bind_rows(
    base_bp %>%
    group_by(DELTA_SBP_GROUP) %>%
    summarise(cnt = length(unique(STUDY_ID)),.groups='drop') %>%
    mutate(
      SITE = 'ALL',
      N = N,
      prop = round(cnt/N,3)
    )
  ) %>%
  select(SITE,DELTA_SBP_GROUP, N, cnt, prop)
elig_drift
```

As shown above, the run-in period duration averaged around `r -round(mean(base_bp$ELIG_SINCE_INDEX))` (`r round(sd(base_bp$ELIG_SINCE_INDEX),1)`) days, and we identified about `r round(elig_drift$prop[1]*100)`% with significant eligibility drift (to below 130 at baseline).

## Investigation: How much of the eligibility drift can be explained by anti-hypertensive medication use during the run-in period?

We set out to investigate what could be some legitimate reasons occured during the run-in period (i.e., the time period between eligibility and recruitment/randomization) to explain this eligibility drift. For this investigation, we also collected all prescription data from CDM PRESCRIBING table provided by KUMC team and identified the following anti-HTN medications (using VA classifications) as well as their timings (start before eligibility visit or during the run-in period):

-   CV100: beta blockers/related

-   CV150: alpha blockers/related

-   CV200: calcium channel blockers

-   CV400: anti-hypertensive combinations

-   CV490: antihypertensive, others

-   CV701: thiazides/related diuretics

-   CV702: loop diuretics

-   CV703: carbonic anhydrase inhibitor diuretics

-   CV704: potassium sparing/combinations diretics

-   CV709: diuretics, other

-   CV800: ACE inhibitors

-   CV805: angiotensin II inhibitor

-   CV806: direct renin inhibitor

### Analysis I: Using systematic explanatory feature screening to investigate what are the explanatory factors for significant BP drop

```{r}
#| label: load-all-data

baseline_aset<-readRDS(file.path(path_to_data,'baseline_aset.rda'))

var_excld<-c(
  "STUDY_ID",
  "DELTA_SBP",
  "DELTA_SBP_GROUP",
  "decr_ind",
  "sig_decr_ind1",
  "sig_decr_ind2",
  "RX_TIMING",
  "BSBP",
  "BSBP_GROUP"
)
var_lst<-colnames(baseline_aset)[!colnames(baseline_aset) %in% var_excld]
var_lst<-var_lst[!grepl("(bef_1st)+?$",var_lst)]
```

Using prescription data, we included `r length(var_lst)` factors that includes:

-   `r length(var_lst[grepl('_maintain',var_lst)])` anti-hyptertensive drug use pattern (generic name & class) indicators suggesting a drug started before and continued during the run-in period;
-   `r length(var_lst[grepl('_runin_start',var_lst)])` anti-hyptertensive drug use pattern (generic name & class) indicators suggesting a drug started during the run-in period;
-   `r length(var_lst[grepl('_runin_stop',var_lst)])` anti-hyptertensive drug use pattern (generic name & class) indicators suggesting a drug started before but no new prescription was observed during the run-in period;
-   `r length(var_lst[grepl('_does_inc',var_lst)])` anti-hyptertensive drug use pattern indicators suggesting a dose increase of a generic drug
-   age at baseline, sex, race, ethnicity, eligibility SBP

```{r}
#| label: med-use-segment

# CreateTableOne(
#   vars = c("DELTA_SBP_GROUP"),
#   strata = "antihtn_delta_med" ,
#   data = baseline_aset,
#   factorVars = c("DELTA_SBP_GROUP")
# )

CreateTableOne(
  vars = c("antihtn_delta_med"), 
  strata = "DELTA_SBP_GROUP" , 
  data = baseline_aset, 
  factorVars = c("antihtn_delta_med")
)
```

```{r}
#| label: med-use-segment2

# CreateTableOne(
#   vars = c("DELTA_SBP_GROUP"), 
#   strata = "antihtn_dose_inc" , 
#   data = baseline_aset, 
#   factorVars = c("DELTA_SBP_GROUP")
# )

CreateTableOne(
  vars = c("antihtn_dose_inc"), 
  strata = "DELTA_SBP_GROUP" , 
  data = baseline_aset, 
  factorVars = c("antihtn_dose_inc")
)
```


```{r}
#| label: simple-fs-multi-re

var_excld<-c(
  "STUDY_ID",
  "DELTA_SBP",
  "DELTA_SBP_sign",
  "DELTA_SBP",
  "decr_ind",
  "sig_decr_ind1",
  "sig_decr_ind2",
  "RX_TIMING",
  "INDEX_DATE"
)
var_lst<-colnames(baseline_aset)[!colnames(baseline_aset) %in% var_excld]
var_lst<-var_lst[!grepl("(bef_1st)+?$",var_lst)]

odds_ratio<-data.frame(
  var1=as.character(),
  var2=as.character(),
  odds_ratio=as.numeric(),
  ci_lower=as.numeric(),
  ci_upper=as.numeric(),
  p_value=as.numeric()
)

for (vari in var_lst){
  cat("test variable:",vari,"\n")
  
  fit<-glm(
    as.formula(paste0("sig_decr_ind1 ~",vari)),
    data=baseline_aset,
    family=binomial()
  )
  summ.fit<-summary(fit)
  ci.fit<-confint(fit)
  
  odds_ratio<-odds_ratio %>%
    add_row(
      var1 = vari,
      var2 = attr(summ.fit$coefficients,"dimnames")[[1]][-1],
      odds_ratio=exp(summ.fit$coefficients[-1,1]), # note that the coefficient of logistic regression is the log-odds
      ci_lower = exp(ci.fit[-1,1]),
      ci_upper = exp(ci.fit[-1,2]),
      p_value=summ.fit$coefficients[-1,4])
}
odds_ratio<-odds_ratio %>% 
  filter(p_value<0.1) %>%
  arrange(p_value)

odds_ratio

var_sel<-unique(c(
  'SITE',
  odds_ratio$var1
))
facvar_sel<-var_sel
cohort_summ<-univar_analysis_mixed(
  df = baseline_aset,
  id_col ="STUDY_ID",
  var_lst = var_sel,
  grp = baseline_aset$sig_decr_ind1,
  facvar_lst  = facvar_sel,
  pretty = T
)
cohort_summ %>% kable_classic() 
```

```{r}
#| label: multi-reg
#| warning: false
#| eval: false

# multiple regression model
var_sel<-odds_ratio$var1
sel_model <- glm(
  formula(paste0("sig_decr_ind1 ~ ",paste(var_sel,collapse = "+"))), 
  data = baseline_aset, family = binomial()
)
sel_model %>% tbl_regression(exponentiate = TRUE)

# pROC::ggroc(pROC::roc(sel_model$y, sel_model$fitted))+
#   geom_abline(intercept=1,linetype=2)+
#   labs(subtitle = paste0("AUC:",round(pROC::auc(sel_model$y, sel_model$fitted),4)))
```

```{r}
#| label: stepwise-reg

# Specify a null model with no predictors
null_model <- glm(
  sig_decr_ind1 ~ 1,
  data = baseline_aset, 
  family = binomial()
)

# Specify the full model using all of the potential predictors
full_model <- glm(
  formula(paste0("sig_decr_ind1 ~ ",paste(var_sel,collapse = "+"))), 
  data = baseline_aset, 
  family = binomial()
)

# Use a bi-directional stepwise algorithm to build a parsimonious model
step_model <- step(
  null_model,
  scope = list(lower = null_model, upper = full_model),
  direction = "both"
)

modsumm<-summary(step_model)
with(modsumm, 1 - deviance/null.deviance)

step_model %>% tbl_regression(exponentiate = TRUE)

# pROC::ggroc(pROC::roc(step_model$y, step_model$fitted))+
#   geom_abline(intercept=1,linetype=2)+
#   labs(subtitle = paste0("AUC:",round(pROC::auc(step_model$y, step_model$fitted),4)))

```

### Analysis II: Profile the affected cohort with reasonable explanations

```{r}
#| label: profile-suseptible-cohort
#| warning: false

sel_aset<-data.frame(baseline_aset) %>% 
  filter(sig_decr_ind1 == 1) %>%
  mutate(
    expln1 = as.numeric(ESBP_GROUP=="esbp1"|antihtn_dose_inc==1|antihtn_add==1|antihtn_runin==1)
  )

var_expln<-c(
  "ESBP_GROUP",
  "antihtn_dose_inc",
  "antihtn_add",
  "antihtn_runin",
  "antihtn_mt",
  "untreated",
  "expln1"
)
  
cohort_summ<-univar_analysis_mixed(
  df = sel_aset,
  id_col ="STUDY_ID",
  var_lst = var_expln,
  facvar_lst  = var_expln,
  pretty = T
)
cohort_summ %>% kable_classic()
```

As shown in the table above, anti-hypertensive drug use and low eligibility SBP together only explains 64% of the eligibility drift.

Historical BP Fluctuation
